\chapter{Assessment Procedure}

The objective of the assessment of the project is to determine and appraise your ability to:

\begin{itemize}
    \item Diagnose and structure a problem
    \item Apply techniques and concepts from the course in a meaningful way
    \item Present logical results and conclusions deriving from analysis
    \item Express any limitations with the method used
    \item Demonstrate an understanding of problems involved with the implementation of proposed solutions
    \item Communicate both in writing and orally, clearly and concisely
    \item Present professional and well formatted documentation
\end{itemize}

Project elements will be marked in accordance with the university assessment regulations. Project reports may be sent to external examiners, and the external examiners may visit to assess the general progress of the projects. 

\begin{tcolorbox}
    Normal University regulations will apply for late submissions. It doesn't matter that the project is worth a lot of your degree classification - keep that in mind and \textbf{do not submit late!}
\end{tcolorbox}

\section{Assessment criteria}


As with all assessments, you will be given an outline of the criteria used by the assessors so that everyone has a clear idea about which aspects of the project and the report are considered to be the most important. These are shown in the assessment briefs which are available on SOL. Please read them carefully alongside this document before starting work on your project.


\subsection{Second marking \& moderation procedure for project reports}

Due to the fact that the project report is responsible for such a large proportion of your degree classification, there is a more stringent marking procedure in place.

All project final reports are fully blind second marked (another assessor is assigned and the project report is marked as though it were a fresh unmarked report, and the other assessor does not know the grade).

If the marks awarded by the first and second markers differ, the following protocol is in place:

\begin{itemize}
    \item If the grades differ by one grademark (eg. 68 and 65) the higher grade will be recorded (eg. 68)
    \item If the grades differ by two grademarks (eg. 68 and 62) then the middle grade will be recorded (eg. 65)
    \item If the grades differ by more than two grademarks (eg. 68 and 58) then the second markers must meet to discuss the project and agree a mark.
    \item If the markers cannot agree on a mark, a third marker will be appointed.
\end{itemize}

\subsection{Definition \& Poster presentation}

Definition marks will be awarded upon the supervisor's grade only. The usual university regulations apply regarding moderation, but it will not be second marked.

All poster presentations will be second marked as per the procedure in the section above.